\documentclass[12pt,a4]{article}
\usepackage[top=.75in, bottom=.75in, left=.75in, right=.75in]{geometry}
\usepackage{titlesec}

\titleformat*{\subsection}{\large\it\bfseries}
\titlespacing*{\section}{0pt}{6ex}{0ex}

\begin{document}

\noindent{\LARGE\textbf{Ethical Scenarios}}

\vspace{1em} 
\noindent Imagine yourself as the researcher in each of the following scenarios. What is ethical, responsible, and methodologically necessary in each situation? {\em Note: There may not be a right answer.}

\begin{enumerate}\itemsep1em

\item The researcher is interested in how the experience of pain affects individuals' physiological responses to threatening stimuli. To study this, the researcher randomly assigns individuals to a painful experience (holding their bare hand in a bowl of ice water for 10 minutes) or a control condition before exposing them to videos containing images of warfare and violence. % Benefit/harm trade-offs

\item The researcher wants to know if mosquito nets help reduce the spread of malaria in a Sub-Saharan Africa country, which prior research suggests are effective but the size of the effect is ambiguous. Some regions in the country have many malaria-carrying mosquitos, while some regions have few. The researcher randomly assigns households across all regions to receive mosquito nets or not. % ethics of randomization

\item The researcher is interested in whether there is gender discrimination in hiring processes. In particular, the researcher thinks that men are discriminated against in hiring. The researcher obtains access to the HR records for several firms, codes each job applicant's qualifications on a number of dimensions, and regresses an indicator for whether the applicant was (1) interviewed and (2) hired against those covariates. The researcher finds anti-male discrimination, controlling for other factors.

\item The researcher is interested in whether giving grants to local libraries for community outreach initiatives increases use of the libraries. Libraries are randomly assigned to receive either no grant, a small grant (5,000 GBP), or a large grant (50,000 GBP). After the study is over, the researcher learns that due to an administrative error, several libraries that were supposed to receive large grants received no grants, and vice versa. % ethical violations

\item All details are the same as the previous scenario, but the ``administrative error'' was made intentionally by the spouse of a local library administrator.

\item The researcher is interested in whether international study experiences by high school students affects the completion time for their post-secondary degrees (either positively or negatively). One thousand students at a university are invited to participate in the study, which will use a lottery to allocate scholarships for stays abroad. Twenty of the students who do not receive a scholarship in the lottery apply elsewhere for a scholarship and eventually study go abroad. % compensatory behavior

\item The researcher is interested in the relationship between parental use of narcotics and long-term educational attainment. The researcher gains permission to use NHS records on narcotics use and links this to publicly specified educational degrees on individuals' LinkedIn and Facebook profiles. No participants in the study are informed of their role in the study.

\item The researcher is interested in interested in the extent to which individuals racially discriminate against out-groups. To do so, the researcher recruits participants for in-depth, semi-structured interviews with a promise of anonymity. The researcher records the participants names and later discloses the anonymized data along with a list of participants (but the responses cannot be linked to each individual) during a media interview.

\item All details are the same as the previous scenario, but the responses can be linked to individuals.

\item All details are the same as the previous scenario, but the researcher did not promise the respondents anonymity.

\item All details are the same as the previous scenario, but the participants are all bureaucrats working in an authoritarian regime.

\item The researcher is interested in testing a game theoretic expectation about levels of social trust after repeated interpersonal interactions. Individuals randomly two conditions play a ``trust game'' with another participant for ten rounds, but individuals in the treatment condition are surprised by a eleventh round where they can potentially lose funds to the other participant, while those in the control condition proceed directly to an outcome questionnaire. % Deception: Omission

\item The researcher aims to understand how narcotics sales and distributions networks operate within London. To do this, the researcher uses their own personal funds to purchase illicit drugs from different individuals in different places in the city without disclosing their role as a researcher. They then work with a chemist to identify shared and distinct chemical composition of each product purchased from the various dealers, thereby constructing a social network of narcotics distributors based on sales of identical products. % Deception: Commission

\item The researcher's goal is to study the effects of bribe-taking on local governments' responsiveness to citizen concerns. The researcher surveys residents in several municipalities about their concerns, encourages anyone with a concern to contact a local official, to offer a bribe and report back on the results their experience.

\item All details are the same as in the previous scenario, but the researcher pays some of the respondents money to use when offering the bribe.

\item The researcher is interested in the effect of an early childhood intervention on individuals' wages in adulthood and plans to follow a group of children for 30 years. Knowing that the study will be difficult, the researcher plans to follow 500 individuals. After ten years, 80\% of treatment group participants were reinterviewed but only 60\% of control group participants were able to be interviewed. % Attrition

\item All details are the same as in the previous scenario, but after ten years all of participants data are accidentally made public on the internet.

\item The researcher is interested in how responsive government officials are to individuals to immigrant and native-born citizens and suspects that members of parliament are more likely to respond and help the native born population than immigrants. The researcher randomly assigns MPs to receive letters requesting assistance from individuals with either Arabic-sounding names or English-sounding names in order to measure and record their responses. The researcher publishes the results in a major newspaper editorial that criticizes the government. % Informed consent

\item TfL is interested in assessing whether disruptions to Tube service lead people to use buses instead. To study this, they provide a complete dataset of passenger trips (based on Oyster card transactions) and known travel disruptions to a university research student to assess the relationship.

\item The researcher works for a political party and is trying to design effective campaign materials for their party organization to use in the next election. The researcher designs several versions of campaigns slogans and constructs a representative sample of the population. The researcher then places phone calls to each individual describing the survey interview as an independent public opinion poll and exposes participants to some of the slogans and measures their responses. The researcher uses the results of the study to design the party's next election campaign. % informed consent

\item The researcher's goal is to find the most cost-effective means of providing long-term health and social care to prisoners with mental health conditions. The researcher randomly assigns individuals in this subpopulation to live in either a small residential-like setting or a traditional prison setting. The researcher finds the large, hospital-like setting to be the most cost-effective. % Vulnerable population

\item The researcher is interested in studying how religious minority groups think about the majority population in a country. The researcher pretends to be a member of the selected religious group, attends ceremonies and other events, and discusses the research topic with members without disclosing their role as a researcher.

\item All details are the same as in the previous scenario, but the researcher instead discloses their role as a researcher.

\item The researcher is interested in whether ticket enforcement on municipal buses is cost effective (i.e., whether the cost of paying ticket enforcers is outweighed by the gain from more riders paying for their tickets). The researcher believes that ticket violations (i.e., not buying a ticket) are highest in the city's ten districts with the lowest median income. The researcher assigns these districts to receive increased enforcement and measures the change in ticket purchases across all districts over time. %Differential enforcement across regions; generalizability to other regions

\item The researcher wants to know if leadership training for ministerial managers improves the productivity of their employees. The training program's effectiveness will be assessed every six months and continue for three years. After 12 months the program shows a modest, positive effect but it is hard to distinguish from no gain. After 18 months, the effect appears to be smaller but still positive. The ministry receives additional funding for the program and suggests to the researcher that the study be discontinued and the program expanded to all managers. % Discontinuing a study early: apparent effectiveness

\end{enumerate}

\end{document}