\documentclass[a4, 12pt]{article}
\usepackage[top=2cm, bottom=2cm, left=2cm, right=2cm]{geometry}

\title{Theory and hypothesis generation}
\author{}
\date{}

\begin{document}

\maketitle

\maketitle
\vspace{-6em}

\section{Purpose}
The purpose of this assignment is to assess your ability to identify, evaluate, and generate political science theories.

\section{Overview}

You are being given a published article to read and evaluate. For this concept, you need to identify the theoretical argument being presented, generate rival theoretical explanations, evaluate the quality of the published theory and one or more alternatives, and assess the extent to which the article adequately describes and tests the observable implications of its theory.

\section{Your Task}

\begin{enumerate}\itemsep1em
\item Read the article ``The Nuclear Taboo'' by Nina Tannenwald. What is her research question? What does she want to understand?

\item What is Tannenwald's theory? Considering the criteria for evaluating a ``good'' theory, how ``good'' is Tannenwald's theory? How could it be improved?

\item What are one or more alternative explanations for the outcome(s) she is attempting to explain? Why are those alternatives better or worse? How would you empirically distinguish each theory from the other(s), or are they empirically indistinguishable?

\item Overall, how well does Tannenwald articulate and empirically test her theory? What could be improved?

\item In a few sentences, briefly reflect on what was easy or difficult about this assignment and provide any feedback on the assignment for the instructor.

\end{enumerate}

\section{Submission Instructions}

Your assignment should be no more than 2 pages of double-spaced A4, Times New Roman font size 12. Submit the assignment via Moodle by Tuesday, November 10.

\section{Feedback}

You will receive feedback within two weeks.


\end{document}
